---
title: "Interacting with APIs"
author: "Sandra Aziz"
output: 
  github_document:
      toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
This document is an vignette to show how to interact with an API. I will demonstrate this by interacting with the Movie API and perform some exploratory data analysis (EDA) on the data extracted.  
I will create a function that takes in four arguments: the movie title, the release year (defaults to NULL), the plot (defaults to "short"), and the endpoint (defaults to "movie"), and, if the movie is present on the API, the function will return a data frame that includes the title of the movie, the year it was released in, its genre, plot, and IMDb rating.  

# Required Packages
To use the function for interacting with the Movie API, I used the following packages:  
* `httr`: to make HTTP requests  
* `jsonlite`: to read JSON data and convert and simplify to an R object  

In addition to those packages, I used the following packages in the EDA portion:  
* `dplyr`: to manipulate and summarize data  
* `ggplot2`: to visualize data  

```{r 1, warning=FALSE}
# Load the necessary libraries
library(httr)
library(jsonlite)
library(dplyr)
library(ggplot2)
```

# API Interaction Function
After loading the required libraries, I will create the function to call the Movie API:  
```{r 2}
get_movie_data <- function(title, year = NULL, plot = "short", endpoint = "movie") {
  # Prepare the API request URL
  url <- "http://www.omdbapi.com/"
  params <- list(apikey = "335677ef" , t = title, y = year, plot = plot, r = "json", type = endpoint)

  # Send the API request
  response <- httr::GET(url, query = params)

  # Parse the JSON response into a data frame
  data <- jsonlite::fromJSON(content(response, "text"))

  # Return data frame
  if (data$Response == "True") {
     df <- data.frame(
      Title = data$Title,
      Type = data$Type,
      Run_Time = as.numeric(gsub(" min", "", data$Runtime)),
      Year = as.integer(data$Year),
      Box_Office = as.numeric(gsub("[$,]", "", data$BoxOffice)),
      Genre = data$Genre,
      Metascore = data$Metascore,
      IMDb_Rating = as.numeric(data$imdbRating)
    )
    return(df)
  } else {
    stop("Error: Movie not found.")
  }
}
```

# Data Exploration
Now that we can interact with a few API endpoints, let's do some data exploration and analysis!  
First, I will feed the function the titles for popular movies from different decades to get information about them , combine the information into one table, then do some analysis on them.  

```{r 3}
A <- get_movie_data("The Godfather")
B <- get_movie_data("Scarface")
C <- get_movie_data("The Dark Knight")
D <- get_movie_data("The Social Network")
E <- get_movie_data("The Amazing Spider-Man")
G <- get_movie_data("The Hunger Games")
H <- get_movie_data("White House Down")

movies <- rbind(A, B, C, D, E, G, H)

# Create a new variable "Decade" based on the release year
movies$Decade <- floor(movies$Year / 10) * 10

movies
```

Now, I will creating a contingency table of movie genres:  
```{r 4}
genre_table <- table(movies$Genre)
genre_table
```

From the table above, we can tell that 2 out of the 7 movies selected are considered part of the crime and drama categories. However, if we break the genres into their separate categories, 5 out of 7 movies would be considered drama and 3 out of the 7 would be considered crime.  

Next, I will create numerical summaries of rating at each category of genre:
```{r 5}
genre_summaries <- aggregate(movies$IMDb_Rating, by = list(movies$Genre), FUN = summary)
genre_summaries
```
The summaries above show the spread of the IMDb_Rating variable across the different genres. Of course, there isn't any variation for the second, third, and fourth groups of genres since they only include one movie each from our list. The IMDb_Rating for first group ranges from 6.9 to 7.2 and for the last group from 8.3 to 9.2 for the two movies in teach of hese categories. Of course the more movies in each categories, the more variation we expect to see in their IMDb_Rating.  

Next, I will create some plots so we can visualize this better:
```{r 6}
bar_plot <- ggplot(movies, aes(x = Title, y = IMDb_Rating, fill = Genre)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Movie Ratings by Genre", x = "Movie Title", y = "IMDb Rating") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
bar_plot
```
From this bar plot, we can see that the movies in 'Crime, Drama', 'Action, Crime, Drama' and 'Biography, Drama' genres from our list have higher IMDb Rating than the movies in the other genres. It would be interesting to know if this is a trend throughout these genres or just in the movies that were picked.   

```{r 7}
hist_plot <- ggplot(movies, aes(x = IMDb_Rating, fill = Genre)) +
  geom_histogram(binwidth = 0.25) +
  labs(title = "Distribution of Movie Ratings", x = "IMDb Rating", y = "Count")
hist_plot
```
We can tell that the IMDb Rating for the 2 movies in the 'Action, Adventure, Sci-Fi' genre are closely related. This makes me curious if these results are rare or if all the movies in this genre would most likely fall around the same IMDb rating.

```{r 8}
box_plot <- ggplot(movies, aes(x = Genre, y = IMDb_Rating, fill = Genre)) +
  geom_boxplot() +
  labs(title = "Boxplot of IMDb Ratings by Genre", x = "Genre", y = "IMDb Rating") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
box_plot
```
This boxplot visualizes summary table discussed early. Similarly, we can see the three genres with one movie each that have no variation compare to the other two. We can also see how closely the IMDb Rating is for the two movies in the 'Action, Adventure, Sci-Fi' compared to the 'Crime, Drama' genre as discussed in the histogram above.  

```{r 9}
scatter_plot <- ggplot(movies, aes(x = IMDb_Rating, y = Metascore, color = Genre)) +
  geom_point() +
  labs(title = "IMDb Ratings vs. Metascores", x = "IMDb Rating", y = "Metascore") 
scatter_plot
```
If we exclude the 'Crime, Drama' genre, we would be able to see a clear trend 

```{r 10}
genre_ratings <- aggregate(IMDb_Rating ~ Genre, data = movies, FUN = mean)
bar_plot2 <- ggplot(genre_ratings, aes(x = Genre, y = IMDb_Rating, fill = Genre)) +
  geom_bar(stat = "identity") +
  labs(title = "Average IMDb Ratings by Genre", x = "Genre", y = "Average IMDb Rating") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
bar_plot2
```
The plot above shows the average IMDb_Rating for each genre. These numbers can also be found in the numerical summaries discussed above.  

# Wrap-Up
So far, to summarize everything above, I created a function that interacts with the Movie API and takes in the title of the movie, series, or episode, and outputs relevant information about them. Then, I created some categorical, numerical, and graphical summaries that explored more about the movies I picked. All the summaries seemed to agree with each other, which is good to achieve! I am interested in looking into this more by extracting titles of different types and doing some cross analysis at a later time! I hope this has helped you interact with APIs!
```